Make any stupid plot you want

### 2021-01-25 draft/sections/math.tex
fleshing out H

\subsubsection{Variables: Fiber Space $F$}
\label{sec:data_fiber}
\note{yes, I know i may need to change this U to like \mathcal{U}, spivak uses U}
The fiber is a topological space that is the set of possible values of the data; the values themselves can be any dimension and type and have any continuity. We use Spivak's description of simplicial database schemas \cite{spivakSIMPLICIALDATABASES} as the basis of our fiber space because he binds the components of the fiber to variable names and types. Spivak constructs a set $U$ that is the disjoint union of all possible objects of types $\{T_0, \ldots, T_n\} \in DT$ where $DT$ are the data types of the variables in the dataset. He then defines the single variable set $U_\sigma$ 
\begin{equation}
    \label{eq:data_types}
\begin{tikzcd}
    U_{\sigma} \arrow[r] \arrow[d, "\pi_{\sigma}"'] & U \arrow[d, "\pi"] \\
    C \arrow[r, "\sigma"']                          & \textbf{DT}       
\end{tikzcd}
\end{equation}
which is $U$ restricted to objects of type $T$ bound to variable name $c$. Given $U_{\sigma}$, the fiber for a one variable dataset is
\begin{equation}
    F = U_{\sigma(c)} = U_{T} 
\end{equation}
where $\sigma$ is the schema binding variable name $c$ to its datatype $T$. A dataset with multiple variables has a fiber that is the cartesian cross product of $U_{\sigma}$ applied to all the columns:
\begin{equation}
F = U_{T_1}\times \ldots U{T_i} \ldots\times U_{T_n}
\end{equation}

@@

For example, the data in figure~\ref{fig:fiber_example_plane} is a pair of times and \textdegree C temperature measurements taken at those times. Time is a positive number ot type \textt{detetime} which can be resolved to positive floats $U_{\texttt{datetime}}= \mathbb{R}^+$. Temperature values are real numbers $U_{\texttt{float}} = \mathbb{R}$. The fiber is 
\begin{equation}
    F =  \mathbb{R}^{+} \times \mathbb{R} 
\end{equation} 
where $V_0 = (time,\, \texttt{datetime},\, \mathbb{R}^+)$ and $V_1 = (temperature,\, \texttt{float},\, \mathbb{R})$. In figure~\ref{fig:fiber_example_cube}, temperature is replaced with wind. This wind variable is of type \texttt{wind} and has two components speed and direction $\{(s,d) \in \mathbb{R}^{2} \mid  0\leq s,\, 0 \leq d \leq 360\}$ . Therefore, the fiber is 
\begin{equation}
    F = \mathbb{R}^{+} \times \mathbb{R}^2
\end{equation} 
such that $V_1 = (wind,\, \texttt{wind},\, \mathbb{R}^{2})$

@@

Steven's described the measurement scales\cite{stevensTheoryScalesMeasurement1946,leaFormalizationMeasurementScale} in terms of the monoid actions on the measurements: nominal data is permutable, ordinal data is monotonic, interval data is translatable, and ratio data is scalable \cite{weissteinSimilarityTransformation}. For example, given  $V_0 = (temperature,\, \txtett{float},\, \mathbb{R})$ which is interval data:
\begin{itemize}
    \item monoid operator addition $\ast = +$
    \item monoid operations: $f: x\mapsto x + c_1 $, $g: x\mapsto x + c_2$
    \item monoid action operator composition $\bullet = \circ$
\end{itemize}
then the translation monoid actions on temperature satisfy the condition
\begin{equation}
    \begin{tikzcd}
        \mathbb{R} \arrow[rd, "(x+c_1)\circ(x+c_2)"] \arrow[d, "x+c_1"'] &            \\
        \mathbb{R} \arrow[r, "x+c_2"']                                   & \mathbb{R}
    \end{tikzcd}
\end{equation}
where $c_1, c_2$ are intervals $x_i-x_j$.

@@

\subsection{Prerender Space $H$}
\label{sec:graphic}

We define a graphic space $H$ such that we do not have to assume the physical output space of the renderer. This means that the graphic in $H$ can be output to a screen or 3D printed space or a dome. We model the prerender space as a fiber bundle (H, S, $\pi$, D). $H$ is the predisplay space, with a fiber $D$ dependent on the target display and a base space of $S$.

%%define a region b = [top, bottom, right, left]
%% inverse of b is capital S_b
\subsubsection{Base space}
The underlying topology $S$ of a graphic often needs more dimensions than the data topology $K$ because of the specifications of the display space. For example, a line plot on a plane (such as a screen or a piece of paper) by necessity needs to also have a thickness so that it is visible, which maps back to a set of connected points in $H$. The topology of these connected points is therefore the region $s \subset S$ such that $\xi: S \rightarrow K$ is a deformation retraction \cite{RetractionTopology2020}
\begin{equation}
    \begin{tikzcd}
        E \arrow[d, "\pi"'] & H \arrow[d, "\pi"'] \\
        K                   & S \arrow[l, "\xi"']
        \end{tikzcd}
\end{equation}

that goes from a region $s \in S_{k}$ to its associated point $k$, such that when $\xi(s) = k$, $\xi^*\tau(s) = \tau(k)$.

\subsubsection{Fiber and Section}
A section $\rho: S \rightarrow H$ is a mapping from a region $s$ on a mathematical encoding of the image to a region $xy$ on the screen that the renderer then maps to visual space as defined in D.

>>>>

\subsubsection{Variables: Fiber Space $F$}
\label{sec:data_fiber}
\note{yes, I know i may need to change this U to like $\mathcal{U}$, spivak uses U}
The fiber is a topological space that is the set of possible values of the data; the values themselves can be any dimension and type and have any continuity. We use Spivak's description of simplicial database schemas \cite{spivakSIMPLICIALDATABASES} as the basis of our fiber space because he binds the components of the fiber to variable names and types. Spivak constructs a set $U$ that is the disjoint union of all possible objects of types $\{T_0, \ldots, T_n\} \in DT$ where $DT$ are the data types of the variables in the dataset. He then defines the single variable set $U_\sigma$ 
\begin{equation}
    \label{eq:data_types}
\begin{tikzcd}
    U_{\sigma} \arrow[r] \arrow[d, "\pi_{\sigma}"'] & U \arrow[d, "\pi"] \\
    C \arrow[r, "\sigma"']                          & \textbf{DT}       
\end{tikzcd}
\end{equation}
which is $U$ restricted to objects of type $T$ bound to variable name $c$. Given $U_{\sigma}$, the fiber for a one variable dataset is
\begin{equation}
    F = U_{\sigma(c)} = U_{T} 
\end{equation}
where $\sigma$ is the schema binding variable name $c$ to its datatype $T$. A dataset with multiple variables has a fiber that is the cartesian cross product of $U_{\sigma}$ applied to all the columns:
\begin{equation}
F = U_{T_1}\times \ldots U{T_i} \ldots\times U_{T_n}
\end{equation}

@@

For example, the data in figure~\ref{fig:fiber_example_plane} is a pair of times and \textdegree C temperature measurements taken at those times. Time is a positive number ot type \texttt{detetime} which can be resolved to positive floats $U_{\texttt{datetime}}= \mathbb{R}^+$. Temperature values are real numbers $U_{\texttt{float}} = \mathbb{R}$. The fiber is 
\begin{equation}
    F =  \mathbb{R}^{+} \times \mathbb{R} 
\end{equation} 
where $V_0 = (time,\, \texttt{datetime},\, \mathbb{R}^+)$ and $V_1 = (temperature,\, \texttt{float},\, \mathbb{R})$. In figure~\ref{fig:fiber_example_cube}, temperature is replaced with wind. This wind variable is of type \texttt{wind} and has two components speed and direction $\{(s,d) \in \mathbb{R}^{2} \mid  0\leq s,\, 0 \leq d \leq 360\}$ . Therefore, the fiber is 
\begin{equation}
    F = \mathbb{R}^{+} \times \mathbb{R}^2
\end{equation} 
such that $V_1 = (wind,\, \texttt{wind},\, \mathbb{R}^{2})$

@@

Steven's described the measurement scales\cite{stevensTheoryScalesMeasurement1946,leaFormalizationMeasurementScale} in terms of the monoid actions on the measurements: nominal data is permutable, ordinal data is monotonic, interval data is translatable, and ratio data is scalable \cite{weissteinSimilarityTransformation}. For example, given  $V_0 = (temperature,\, \texttt{float},\, \mathbb{R})$ which is interval data:
\begin{itemize}
    \item monoid operator addition $\ast = +$
    \item monoid operations: $f: x\mapsto x + c_1 $, $g: x\mapsto x + c_2$
    \item monoid action operator composition $\bullet = \circ$
\end{itemize}
then the translation monoid actions on temperature satisfy the condition
\begin{equation}
    \begin{tikzcd}
        \mathbb{R} \arrow[rd, "(x+c_1)\circ(x+c_2)"] \arrow[d, "x+c_1"'] &            \\
        \mathbb{R} \arrow[r, "x+c_2"']                                   & \mathbb{R}
    \end{tikzcd}
\end{equation}
where $c_1, c_2$ are intervals $x_i-x_j$.

@@

\subsection{Graphic: $H$}
\label{sec:graphic}  
We can separate the structure of the graphic from the properties of the output format by modeling the space of graphics as a fiber bundle  $(H, S, \pi, D)$. As with data, the fiber bundle is for a class of graphics with shared base space $S$ (~\ref{sec:graphic_fiber}) and fiber $D$ (~\ref{sec:graphic_base}) and the sections $\rho$ (~\ref{sec:graphic_section}) encode the fully specified graphics.

\subsubsection{Display $D$}
\label{sec:graphic_fiber}
The fiber $D$ is specified in terms of the target display space, for example a 2D screen or 3D printer. In this work, we assume a simple 2D opaque image $D=\mathbb{R}^5$ with a record $d \in D$ of the form
\begin{equation}
    d = (x, y, R, G, B)
\end{equation}
such that a graphic only has 2D position and color. To support overplotting and transparency, the fiber would be $D=R^{7}$ such that $d=(x,y,z, R, G, B, A)$,

%%define a region b = [top, bottom, right, left]
%% inverse of b is capital S_b

\subsubsection{Continuity of the Graphic $S$} 
\label{sec:graphic_base}
An assumption of graphical representations is that they match the continuity of the data, for example discrete data is mapped to discrete elements\cite{tufteVisualDisplayQuantitative2001,friendlyBriefHistoryData2008}, but the underlying topology $S$ of a graphic may need more dimensions than the data topology $K$ because of the specifications of the display space. Therefore we define the base space mapping from graphic $S$ to data $K$ 
\begin{equation}
    \begin{tikzcd}
        E \arrow[d, "\pi"'] & H \arrow[d, "\pi"'] \\
        K                   & S \arrow[l, "\xi"']
        \end{tikzcd}
\end{equation}

 as the deformation retraction \cite{RetractionTopology2020} $\xi: S \rightarrow K$ that goes from a region $s \in S_{k}$ to its associated point $k$, such that when $\xi(s) = k$, $\xi^*\tau(s) = \tau(k)$.

\begin{figure}
    \includegraphics[width=1\textwidth]{figures/math/retraction_maps.png}
    \caption{The scatter and line graphic base spaces $S$ are larger than $K$ so that they can encode physical aspects of the glyph, such as shape (a circle) or thickness. The heatmap base $S$ is the same as $K$ because a heatmap is directly mapping the values in the data to the graphic.
    \note{add $\alpha, \beta$ coordinates}}
    \label{fig:graphic_retraction_map}
\end{figure}
Figure~\ref{fig:graphic_retraction_map} illustrates some examples of $\xi$. The base space $S$ is mapping to a in $D$. For example, in the scatter plot, $S$ encodes what the glyph looks like such that each piece of a single glyph maps to a different region on $S$. The line plot $S$ can encode position in the $\alpha$ dimension, but the $\beta$ dimension describes

The  line plot $S$ encodes on a plane  by necessity needs to have a thickness to be visible; therefore we construct a rectangular base space $S$ that mirrors the plane on which we can encode the position in one dimension and thickness (should this pixel be colored?) in the other. The map $xi$ then goes from a rectangle to

\subsubsection{Renderable Glyph $\rho$}
\label{sec:graphic_section}
A section $\rho: S \rightarrow H$ is a mapping from a region $s$ on a mathematical encoding of the image to a region $xy$ on the screen that the renderer then maps to visual space as defined in D.
